---
title: "Build and deploy a stroke prediction model using R"
date: "`r Sys.Date()`"
output: html_document
author: "Neha Simon"
---

# About Data Analysis Report

This RMarkdown file contains the report of the data analysis done for the project on building and deploying a stroke prediction model in R. It contains analysis such as data exploration, summary statistics and building the prediction models. The final report was completed on `r date()`. 

**Data Description:**

According to the World Health Organization (WHO) stroke is the 2nd leading cause of death globally, responsible for approximately 11% of total deaths.

This data set is used to predict whether a patient is likely to get stroke based on the input parameters like gender, age, various diseases, and smoking status. Each row in the data provides relevant information about the patient.


# Task One: Import data and data preprocessing

## Load data and install packages

```{r}
data <- read.csv("healthcare-dataset-stroke-data.csv")
library(tidymodels)
library(tidyverse)
library(workflows)
library(tune)
library(naniar)         # Data structures and functions for plotting of missing values
library(caTools)        # Several basic utility functions
library(ggplot2)        # Data visualisations Using the Grammar of Graphics
library(superheat)      # Generating customizable heatmaps
library(scatterplot3d)  # Plots a three dimensional point cloud
library(ROCR)           # Creating cutoff-parameterized 2D performance curves 
```


## Describe and explore the data

```{r}
summary(data)
```

```{r}
unique(data $ gender)
unique(data $ ever_married) 
unique(data $ Residence_type)
unique(data $ smoking_status)
unique(data $ work_type)
```

## Data Cleaning 
```{r}
clean_data <- replace_with_na(data = data, replace = list(bmi = c("N/A"), smoking_status = c(3))) %>% mutate(bmi = as.numeric(bmi))
summary(clean_data)
```

```{r}
clean_data <- data %>%
  mutate(
    gender = if_else(gender == "Female", 0, 
                     if_else(gender == "Male", 1, 2)), 
    ever_married = if_else(ever_married == "Yes", 1, 0), 
    Residence_type = if_else(Residence_type == "Rural", 0, 1), 
    smoking_status = if_else(smoking_status == "never smoked", 0, 
                             if_else(smoking_status == "formerly smoked", 1, 
                                      if_else(smoking_status == "smokes", 2, 3))),
    work_type = if_else(work_type == "Private", 0, 
                        if_else(work_type == "Self-employed", 1,
                                 if_else(work_type == "Govt_job", 2, 
                                          if_else(work_type == "children", 3, 
                                                   if_else(work_type == "Never worked", 4, 4)))))
  )

head(clean_data)
```

```{r}
clean_data <- replace_with_na(data = clean_data, replace = list(bmi = c("N/A"), smoking_status = c(3))) %>% mutate(bmi = as.numeric(bmi))%>%
filter(!is.na(bmi))
summary(clean_data)

```
```{r}
clean_data <- na.omit(clean_data)




```
### Visualizing the input

#### Heatmap



```{r}
superheat(subset(clean_data, select = -c(stroke)), scale = TRUE, bottom.label.size = 0.5, bottom.label.text.angle = 90, bottom.label.text.size = 3)
```
##### BMI distribution
```{r}
ggplot(clean_data, aes(x = bmi)) + geom_density(color="black", fill="lightblue") + labs(title = "Distribution of BMI") 
```
#### Gender distribution 
```{r}
ggplot(clean_data, aes(x = gender)) + geom_density(color="Black", fill="purple") + labs(title = "Distribution of BMI") 
```
#### age 
```{r}
ggplot(clean_data, aes(x = age, y = hypertension, color = stroke)) + geom_point() + scale_color_gradient(low = "lightblue", high = "red")

```
#### Age and BMI wrt stroke

```{r}
ggplot(clean_data, aes(x = age, y = bmi, color = stroke)) + geom_point() + scale_color_gradient(low = "lightblue", high = "red")
```
```{r}
ggplot(clean_data, aes(x = stroke, y = avg_glucose_level, group = stroke, fill = stroke)) + geom_boxplot()
```
## Logistic Regression
```{r}
set.seed(99)  # Set a seed for reproducible results
split = sample.split(clean_data $ stroke, SplitRatio = 0.7)
train = subset(clean_data, split == TRUE)
test = subset(clean_data, split == FALSE)
logistic_regression_1 = glm(stroke~., data = train, family = 'binomial')
summary(logistic_regression_1)
```
```{r}
logistic_regression_2 = glm(stroke ~ gender + age + hypertension + heart_disease + ever_married + avg_glucose_level + bmi + smoking_status, data = train, family = 'binomial')
summary(logistic_regression_2)
```
```{r}
logistic_regression_2 = glm(stroke ~ age + hypertension + heart_disease + ever_married + avg_glucose_level + bmi + smoking_status, data = train, family = 'binomial')
summary(logistic_regression_2)
```
###### The least significant variable as seen is bmi with a Pr-value of 0.29928. Hence we will remove bmi.

```{r}
logistic_regression_2 = glm(stroke ~ age + hypertension + heart_disease + ever_married + avg_glucose_level + smoking_status, data = train, family = 'binomial')
summary(logistic_regression_2)
```

###### The least significant variable as seen is heart_disease with a Pr-value of 0.44624. Hence we will remove heart_disease.

```{r}
logistic_regression_2 = glm(stroke ~ age + hypertension + ever_married + avg_glucose_level + smoking_status, data = train, family = 'binomial')
summary(logistic_regression_2)
```

###### The least significant variable as seen is ever_married with a Pr-value of 0.14813. Hence we will remove ever_married.

```{r}
logistic_regression_2 = glm(stroke ~ age + hypertension + avg_glucose_level + smoking_status, data = train, family = 'binomial')
summary(logistic_regression_2)
```

###### The least significant variable as seen is smoking_status with a Pr-value of 0.12771. Hence we will remove smoking_status.

```{r}
logistic_regression_2 = glm(stroke ~ age + hypertension + avg_glucose_level, data = train, family = 'binomial')
summary(logistic_regression_2)
```

##### Hence we get the three most significant variables having Pr-values less than 0.05

# Task Two: Build prediction models
```{r}
predict_train = predict(logistic_regression_2, type = 'response')
table(train $ stroke, predict_train>0.2)
```
```{r}
predict_test = predict(logistic_regression_2, newdata = test, type = 'response')
table(test $ stroke, predict_test>0.2)
```

# Task Three: Evaluate and select prediction models

```{r}
(940+15)/nrow(test)
```



# Task Four: Set the prediction model

```{r}
library(vetiver)
library(pins)
```

```{r}
v <- vetiver_model(logistic_regression_2, "stroke prediction model")
model_board <- board_folder("pins-r", versioned = TRUE)
model_board %>% vetiver_pin_write(v)
```

```{r}
# save model
saveRDS(logistic_regression_2, "LR_model.RDS")
```
# Task Five: Deploy the model



```{r}
library(rsconnect)
rsconnect::setAccountInfo(name='nehasimon',
			  token='3A883CCD4ABB395EDD0200DFB4D5BAD2',
			  secret='ESqnrSESWxEZkyCX+Z7JoWw0dNJSOans5LNOIQw7')
rsconn1ect::deployApp('~/Projects/stroke-prediction/deploy')
```

























